---
title: "bw-exam1"
format: html
---
##Grand research question: What does food consumption in each category look like by country?

#load the packages here. if they are not installed, install in console. 
```{r}
library(tidytuesdayR)
library(tidyverse)
library(leaflet)
```
#Get data 
```{r}
#get data from the one specific date that we want to be able to answer the question
tuesdata<-tt_load('2020-02-18')
#pull food consumption data specifically
fc<-tuesdata$food_consumption
```

#Understand the data 
```{r}
head(fc) #first 6 rows
summary(fc) #summary stats
str(fc) #structure of data
dim(fc) #dimensions (number of rowns and columns)
```

#Understand the variables individually
```{r}
#country -- spatial 
#create small data set with some countries to place markers on map
  exam_1 <- tribble(
  ~place, ~long, ~lat, 
  "Argentina", 38.4161, 63.6167, 
  "Austrailia", 25.2744, 133.7751, 
  "Liberia", 6.4281, 9.4295, 
  "Bangladesh", 23.6850, 90.3563 
)

#create map
leaflet(data = exam_1) |> 
  addTiles() |> 
  addMarkers()

```
```{r}
#food category
#|fig-height: 22
#|fig-width: 22

ggplot(fc, aes(x = food_category)) +
  geom_bar()
```
```{r}
#consumption
ggplot(fc, aes(y = consumption)) +
  geom_boxplot()
```
#Understand overall food consumption 
```{r}
#overall food consumption and food category
ggplot(fc, aes(y = consumption, x = food_category)) +
  geom_boxplot()
```


```{r}
#overall food consumption and country
ggplot(fc, aes(y = consumption, x = country)) +
  geom_boxplot()
```

#Answer grand research question, and add 5 concepts of effective visulizations.

```{r}
#The challenge with this plot is that there are too many countries. It would be helpful to add consumption as the y-axis and then have it stack consumption by country. 
#| fig-cap: "Bar chart of food category, consumption, and faceted by country."
#| fig-alt: "Mulitiple bar charts -- one for each country. The x axis is food category and they are proportional to consumption of each food category."
ggplot(fc, aes(x = food_category, fill = consumption)) +
  geom_bar(position="fill")+
  facet_wrap(~country)+
  scale_fill_viridis_c()  +
  labs(x = "Food Category", y = "Count", title = "Consumption of Food Category by Country")  
```

```{r}
#The challenge with this plot is that there are too many countries for a density plot. 
#| fig-cap: "Density plots of food consumption by country faceted by food category."
#| fig-alt: "Multiple density plots -- one for each food category. The x axis is consumption and the y axis is density. The density plots are overlapped by country. "
ggplot(fc, aes(x = consumption, fill = country)) + 
  geom_density(alpha = 0.5)+
  facet_wrap(~food_category)+
  scale_fill_viridis_d()  +
  labs(x = "Consumption", y = "Density", title = "Consumption of Food Categories in Different Countries")  
```

#As of now, none of these are appropriate because there are too many obsevrations. But the most approproate visulization is the one that gives you a clear view of your outcome variable. 

