---
title: "bw-foodConsumption2"
format: html
---

#Load packages 
```{r}
library(tidytuesdayR)
library(tidyverse)
library(rnaturalearth)
library(sf)
```

#Load data
```{r}
tuesdata <-tt_load('2020-02-18')
fc <- tuesdata$food_consumption
```


##Research question 1: What are the top 5 countries by food consumption?
#Create sub data
```{r}
question1 <- fc |>
  group_by(country) |>
  summarise(avg_consumption = mean(consumption))|>
  arrange(desc(avg_consumption))|>
  slice_head(n = 5)
```

#Create Viz
```{r}
#| fig-cap: "The top 5 countries that consume the most food (all food groups) are Albania, Finland, Lithuania, Netherlands, and Sweden."
#| fig-alt: "This plot shows the average consumption of all food groups. The top 5 countries that consume the most food (all food groups) are Albania, Finland, Lithuania, Netherlands, and Sweden."
ggplot(question1, aes(x = country, y = avg_consumption)) + 
  geom_col()+
  labs(x = "Country", y = "Average Consumption of all Food Groups", title = "The Top 5 Countries that Consume the Most Food")  
```
#Answer to research question: The top 5 countries that consume the most food (all food groups) are Albania, Finland, Lithuania, Netherlands, and Sweden. 

##Research question 2: What are the top 5 countries that consume each food group?
#Create sub data
```{r}
question2 <- fc |>
  group_by(food_category) |>
  arrange(consumption) |>
  slice_max(consumption, n =  5)
```


#Create Viz
```{r}
#| fig-cap: "Average Consumption of Food Group."
#| fig-alt: "This is a faceted bar plot with each facet being a different food group. The bars in each facet/graph represent the top 5 countries that have the highest consumption of each food group. It is hard to tell what countries are the top 5 consumers of each food group with elements of this graph."

ggplot(question2, aes(x = country, y = consumption)) + 
  geom_col()+
  facet_wrap(~food_category)+
  labs(x = "Country", y = "Average Consumption of Food Group", title = "The Top 5 Countries that Consume Each Food Group") 

```
#Answer to research question: It is hard to tell with elements of this graph. 


##Research Question 3
#What does the consumption of each food look like among countries?
```{r}
#| fig-cap: "Consumption of Each Food Group Across the World."
#| fig-alt: "This is a faceted map with each facet being a different food group. Each facet contains a world map that shows how much that food group is consumed."

ne_countries(returnclass = "sf") |>
  select(name, geometry) |> 
  mutate(name = ifelse(name == "United States of America", "USA", name))|>
  mutate(name = ifelse(name == "Bosnia and Herz.", "Bosnia and Herzegovina", name))|>
  mutate(name = ifelse(name == "Czechia", "Czech Republic", name)) |>
  mutate(name = ifelse(name == "Taiwan", "Taiwan. ROC", name)) |>
  left_join(fc |> select (-co2_emmission),
            join_by(name == country)) |>
  pivot_wider(names_from = food_category,
              values_from = consumption) |>
  select(-"NA")|>
  pivot_longer(cols = c(-name, -geometry),
               names_to = "food_category",
               values_to = "consumption") |>
ggplot() +
geom_sf(aes(fill = consumption)) + 
facet_wrap(~food_category) +
theme(legend.position = "bottom") +
labs(title = "Consumption of Each Food Group Across the World") +
scale_color_viridis_d() 
              
  
  
```



