---
title: "Bivariate Viz"
---

Use this file for practice with the **bivariate viz** in-class activity.  Refer to the class website for details.

```{r}
library(ggplot2)
```


```{r}
# Load data
elections <- read.csv("https://mac-stat.github.io/data/election_2020_county.csv")

# Check it out
head(elections)
```
##Exercise 0: Review

#Part a
How many, or roughly what percent, of the 3000+ counties did the Republican candidate win in 2020?

Take a guess.  
  *43%
  
Then make a plot of the winner variable.
```{r}
ggplot(elections, aes(x = winner_20)) +
  geom_bar() +
  labs(x = "winner in 2020")
```

Then discuss what follow-up questions you might have (and that our data might help us answer).
 * A follow up question I have is which counties voted republican (and how many?)
 
##Exercise 1: Quantitative vs Quantitative Intuition Check
```{r}
ggplot(elections, aes(x = repub_pct_16, y = repub_pct_20)) +
  geom_point() 

```


#Exercise 2: 2 Quantitiative Variables
```{r}
# Set up the plotting frame
# How does this differ than the frame for our histogram of repub_pct_20 alone?
# This differs because now there are two categorical variables.
ggplot(elections, aes(y = repub_pct_20, x = repub_pct_16))
```

```{r}
# Add a layer of points for each county
# Take note of the geom!
ggplot(elections, aes(y = repub_pct_20, x = repub_pct_16)) +
  geom_point()


```
```{r}
# Change the shape of the points
# What happens if you change the shape to another number? 
# Changing the number changes the shapes of the data points
ggplot(elections, aes(y = repub_pct_20, x = repub_pct_16)) +
  geom_point(shape = 5)
```
```{r}
# YOU TRY: Modify the code to make the points "orange"
# NOTE: Try to anticipate if "color" or "fill" will be useful here. Then try both.
ggplot(elections, aes(y = repub_pct_20, x = repub_pct_16)) +
  geom_point(color = "orange")
```
 

```{r}
# Add a layer that represents each county by the state it's in
# Take note of the geom and the info it needs to run!
ggplot(elections, aes(y = repub_pct_20, x = repub_pct_16)) +
  geom_text(aes(label = state_abbr))
```

##Exercise 3: Reflect
  *There is a strong postive relationship between republicican support in 2016 and the same in 2020. 
  
##Exercise 4: Visualizing trend 
