---
title: "Mulivariate Viz"
---

Use this file for practice with the **mulivariate viz** in-class activity.  Refer to the class website for details.

```{r}
# Import and check out data
education <- read.csv("https://mac-stat.github.io/data/sat.csv")
head(education)
```
##Exercise 1: SAT scores
#Part a: Construct a plot of how the average sat scores vary from state to state. (Just use 1 variable – sat not State!)
```{r}
ggplot(education, aes(x = sat)) +
  geom_histogram()  
```
#Part b: Summarize your observations from the plot. Comment on the basics: range, typical outcomes, shape. (Any theories about what might explain this non-normal shape?)
```{r}
#scores around 900 are the most common. 
#scores on the extremes are the least common
```

##Exercise 2: SAT Scores vs Per Pupil Spending & SAT Scores vs Salaries
#Part a
```{r}
# Construct a plot of sat vs expend
# Include a "best fit linear regression model" (HINT: method = "lm")

ggplot(education, aes(y = sat, x = expend)) +
  geom_point() + 
geom_smooth(method = "lm", se = FALSE) 
```
#Part b: What are the relationship trends between SAT scores and spending? Is there anything that surprises you?
```{r}
#I am suprised that score goes down as expenditure increases.
```

##Exercise 3: SAT Scores vs Per Pupil Spending and Teacher Salaries
#Construct one visualization of the relationship of sat with salary and expend. HINT: Start with just 2 variables and tweak that code to add the third variable. Try out a few things!
```{r}
#expend and SAT
ggplot(education, aes(y = sat, x = expend)) +
  geom_point() 

#expend, SAT, and salary
ggplot(education, aes(y = sat, x = expend, size = salary)) +
  geom_point() 
```
##Exercise 4: Another way to Incorporate Scale
#It can be tough to distinguish color scales and size scales for quantitative variables. Another option is to discretize a quantitative variable, or basically cut it up into categories.
#Construct the plot below. Check out the code and think about what’s happening here. What happens if you change “2” to “3”?
```{r}
ggplot(education, aes(y = sat, x = salary, color = cut(expend, 2))) + 
  geom_point() + 
  geom_smooth(se = FALSE, method = "lm")

#When we change 2 to 3, the data gets cut into 3 different sections instead of 2
```

##Exercise 5: Finally an Explanation
#Part a: Build a univariate viz of fracCat to better understand how many states fall into each category, fracCat. 
```{r}
ggplot(education, aes(x = fracCat )) +
  geom_bar()  
```

#Part b: Build 2 bivariate visualizations that demonstrate the relationship between sat and fracCat. What story does your graphic tell and why does this make contextual sense?
```{r}
#in both, fracCat varies by color
#histogram
ggplot(education, aes(x = sat, fill = fracCat)) + 
  geom_histogram(position = "fill")


#density 
ggplot(education, aes(x = sat, fill = fracCat)) +
  geom_density() +
  scale_fill_manual(values = c("blue", "purple", "red"))
```

